Some examples in extended logic counting classes

### 2016-04-20 tex/below.tex
19 abr

>>>>

\begin{theo}
	$\XE{2} \subseteq \totp$
\end{theo}
\begin{proof}
	Let $f:\Truc\to\N$ be a function in $\XE{2}$. Let $h$ be the function described in Theorem \ref{sigma2hard}. Consider a non-deterministic $M$ machine that on input $\A$ first computes $\varphi = h(\A)$ and then simulates the $\totp$ machine that computes $\shdhsat$. Since $h$ is a parsimonious reduction from $f$ to $\shdhsat$, it holds that $f(\A) = \tot_M(\A)+1$.
\end{proof}

### 2016-04-20 tex/horn_h.tex
19 abr

We define the decision problem
\[
\textsc{Disj-Horn-Sat} = \{\Phi \mid \Phi \mbox{ is a disjunction of Horn formulas and $\Phi$ is satisfiable}\},
\]
and the counting problem {\sc \#Disj-Horn-Sat} as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas.

\begin{theo}
	$\textsc{Disj-Horn-Sat}$ is hard for $\XE{2}$ under parsimonious reductions. 
\end{theo} 
\begin{proof}
	Let $f$ be an arbitrary function in $\XE{2}$ and let $f_{\textsc{Disj-Horn-Sat}}$ be the function that characterizes {\sc \#Disj-Horn-Sat}. We will now show a function $h:\Truc\to L(P)$ such that for every $\A\in\Truc$, it holds that $f(\A) = f_{\textsc{Disj-Horn-Sat}}(h(\A))$, and which can be computed in polynomial time. Let $\psi$ be such that $f = f_{\psi}$. We separate the proof in three cases:\begin{enumerate}
		\item All free variables in $\psi$ are of first order. Let $\psi = \exists \x \forall \y\,\varphi(\x,\y,\z)$ where $\z = (z_1,\ldots,z_d)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a propositional formula $\Phi = h(\A)$. First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\forall \y\,\varphi(\x,\y,\e)$. Let $n$ be the number of such assignments to $\z$. Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$. The Disj-Horn-Sat formula we return is
		\[
		\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
		\]
		which has exactly $n$ possible assignments. Note that each $p_i$ and $\neg p_i$ is a Horn clause itself.

>>>>

We define the decision problem
\[
\shdhsat = \{\Phi \mid \Phi \mbox{ is a disjunction of Horn formulas and $\Phi$ is satisfiable}\},
\]
and the counting problem {\shdhsat } as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas.

\begin{theo} \label{sigma2hard}
	$\shdhsat$ is hard for $\XE{2}$ under parsimonious reductions. 
\end{theo} 
\begin{proof}
	Let $f$ be an arbitrary function in $\XE{2}$. We will now show a function $h:\Truc\to L(P)$ such that for every $\A\in\Truc$, it holds that $f(\A) = \shdhsat(h(\A))$, and which can be computed in polynomial time. Let $\psi$ be such that $f = f_{\psi}$. We separate the proof in three cases:\begin{enumerate}
		\item All free variables in $\psi$ are of first order. Let $\psi = \exists \x \forall \y\,\varphi(\x,\y,\z)$ where $\z = (z_1,\ldots,z_d)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a propositional formula $\Phi = h(\A)$. First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\forall \y\,\varphi(\x,\y,\e)$. Let $n$ be the number of such assignments to $\z$. Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$. The disjunction of Horn formulas we return is
		\[
		\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
		\]
		which has exactly $n$ possible assignments. Note that each $p_i$ and $\neg p_i$ is a Horn clause itself.

### 2016-04-20 tex/pe.tex
19 abr

@@

\subsection{Counting the number of paths between two given nodes in a graph}

>>>>

In this section we will show some problems that are known to be in $\PE$, a formula that captures the function, if known, and proof that it belongs to $\totp$.

@@

\subsection{\#PATH}

We define $\textsc{\#PATH}$ as a function such that for each non-directed graph $G$, a pair of its nodes $a,b$ and $k\in\N$, it holds that $\textsc{\#PATH}(G,a,b,k)$ is equal to the numbers of paths from $a$ to $b$ with size at most $k$.

\begin{theo}
	 $\textsc{\#PATH} \in \shl$
\end{theo}
\begin{proof}
	Let $M$ be a non-deterministic machine that on input $(G,a,b,k)$ does the following: Start on $a$. On each cycle let $u$ be the previous vertex, and $M$ guesses the next vertex $v$, if $(u,v)\not\in E$, $M$ rejects. Else, $v = u$ and this ends the cycle. If it finds $b$, it accepts. If $k$ cycles have been made without finding $b$, it rejects.

	Clearly, the number of accepting paths of $M$ with input $(G,a,b,k)$ is equal to $\textsc{\#PATH}(G,a,b,k)$. It can be seen that $M$ uses logarithmic space since at each point of the computation it only needs the value of $u$ and a cycle counter.
\end{proof}

\begin{coro}
	 $\textsc{\#PATH} \in \totp$
\end{coro}

\subsection{\#SIMPLEPATH}

We define $\textsc{\#SIMPLEPATH}$ as a function such that $\textsc{\#SIMPLEPATH}(G,a,b)$ counts the number of paths from $a$ to $b$ such that no vertex appears twice or more in the path.

\begin{coro}
	$\textsc{\#SIMPLEPATH} \in \totp$
\end{coro}
\begin{proof}
	Let $M$ be a non-deterministic machine that on input $(G,a,b)$ does the following: If there is no path from $a$ to $b$, stop. Else, choose $i\in\{0,1\}$ if $i = 0$ stop. Else, start on $a$ and save the path $P=\emptyset$. On each cycle let $S = \emptyset$, let $u$ be the previous vertex and $P$ the current path. If $u = b$, stop. Else for each vertex $v$ such that $(u,v)\in E$, check whether there exists a path $Q$ from $v$ to $b$ such that no vertex of $Q$ appears in $P$. If it exists, add $v$ to $S$. Choose some vertex $v'\in S$, add it to $P$ and do $u = v'$. This ends the cycle.

	Clearly the number of paths of $M$ on input $(G,a,b)$ is equal to $\textsc{\#SIMPLEPATH}(G,a,b)+1$. It can also be seen that $M$ runs on polynomial time.
\end{proof}
